Inferring cosmology from gravitational waves using non-parametric detector-frame mass distribution

### 2024-12-16 src/tex/ms.tex
[showyourwork] overleaf sync

The method works best when the intrinsic distribution is well understood.
However, the features present in the currently inferred mass distribution are not so well characterized yet, and their astrophysical origin is a debated topic \citep{Zevin:2017evb, Mapelli:2020vfa, Zevin:2020gbd, Mandel:2018hfr, Marchant:2023wno}.
In this case, the spectral-siren method depends on the choice of the population model, which can introduce biases in the inference \citep{Mastrogiovanni:2021wsd, Mukherjee:2021rtw, Pierra:2023deu}.
Furthermore, the intrinsic mass distribution is degenerate with cosmology.
For example, an intrinsic mass distribution with a linear redshift evolution is degenerate with $H_0$ at low redshift, as they can both redshift the observed distribution in the same way.
More complex redshift evolution can be degenerate with $H_0$ at high redshift.
While there is no reason to think a realistic intrinsic mass distribution will evolve linearly with redshift, this degeneracy still affects the inference of $H_0$.
Due to such degeneracies, astrophysics and cosmology are tightly linked to each other, and proper inference of the astrophysical and cosmological parameters should be done simultaneously, which makes the method computationally expensive.

@@

The consistency between the simulated and inferred parameters is expected, as the mock data is generated with the same population model and cosmological model we used in the analysis.
We can also see that the uncertainty in $H_0$ is larger compared to the analysis in Sec.~\ref{sec:inference_H0}, as more parameters are inferred simultaneously.
The uncertainty in $\sigma$ is relatively large, this may be due to the large uncertainty in the reconstruction at the gap between the low-mass and high-mass peaks, which can be seen in 
Fig.~\ref{fig:simulation_reconstruction}.

@@

We then perform the remaining steps similar to Sec.~\ref{sec:inference_multi}.
We set the bounds for the power-law index of the redshift population model $\kappa$ to $[-100, 100]$, and the bounds for other free parameters are the same as in Table~\ref{tab:bounds}.
The result of the remapping is shown in Fig.~\ref{fig:real_result_multi}, and it is consistent with the posterior distribution reported in \cite{LIGOScientific:2021aug}.
\begin{figure}
    \script{plot_real_result_multi.py}
    \includegraphics[width=\linewidth]{figures/real_result_multi.pdf}
    \caption{
        Optimized parameter distribution for the real data inference.
    }
    \label{fig:real_result_multi}
\end{figure}
The result implies that \variable{output/real_H0.txt}, with symmetric $90\%$-credible intervals around the median.
The uncertainty is slightly smaller than the one reported in \cite{LIGOScientific:2021aug} ($H_0 = 50^{+37}_{-30} \, \mathrm{km/s/Mpc}$), mainly because it only includes the uncertainty from the reconstruction, which underestimates the true uncertainty which also includes the mismatch between the population model and the observed population distribution.
Figure \ref{fig:real_result_multi} shows a peak in the marginalized posterior distribution of $H_0$ around $140\, \mathrm{km/s/Mpc}$.
This means a significant fraction of the reconstructed observed population distribution has an optimized $H_0$ around $140\, \mathrm{km/s/Mpc}$, which again does not include the uncertainty from the mismatch between the population model and the observed population distribution.

>>>>

The method works best when the intrinsic distribution is well understood.
However, the features present in the currently inferred mass distribution are not so well characterized yet, and their astrophysical origin is a debated topic \citep{Zevin:2017evb, Mapelli:2020vfa, Zevin:2020gbd, Mandel:2018hfr, Marchant:2023wno}.
In this case, the spectral-siren method depends on the choice of the population model, which can introduce biases in the inference \citep{Mukherjee:2021rtw, Mastrogiovanni:2021wsd, Pierra:2023deu, LIGOScientific:2020kqk, KAGRA:2021duu, LIGOScientific:2021aug}.
Furthermore, the intrinsic mass distribution is degenerate with cosmology.
For example, an intrinsic mass distribution with a linear redshift evolution is degenerate with $H_0$ at low redshift, as they can both redshift the observed distribution in the same way.
More complex redshift evolution can be degenerate with $H_0$ at high redshift.
While there is no reason to think a realistic intrinsic mass distribution will evolve linearly with redshift, this degeneracy still affects the inference of $H_0$.
Due to such degeneracies, astrophysics and cosmology are tightly linked to each other, and proper inference of the astrophysical and cosmological parameters should be done simultaneously, which makes the method computationally expensive.

@@

The consistency between the simulated and inferred parameters is expected, as the mock data is generated with the same population model and cosmological model we used in the analysis.
We can also see that the uncertainty in $H_0$ is larger compared to the analysis in Sec.~\ref{sec:inference_H0}, as more parameters are inferred simultaneously.

@@

We then perform the remaining steps similar to Sec.~\ref{sec:inference_multi}.
We set the bounds for the power-law index of the redshift population model $\kappa$ to $[-100, 100]$, and the bounds for other free parameters are the same as in Table~\ref{tab:bounds}.
The result of the remapping is shown in Fig.~\ref{fig:real_result_multi}, and it is consistent with the posterior distribution reported in \cite{LIGOScientific:2021aug}.
\begin{figure}
    \script{plot_real_result_multi.py}
    \includegraphics[width=\linewidth]{figures/real_result_multi.pdf}
    \caption{
        Optimized parameter distribution for the real data inference.
    }
    \label{fig:real_result_multi}
\end{figure}
The result implies that \variable{output/real_H0.txt}, with symmetric $90\%$-credible intervals around the median.
The uncertainty is larger than the one reported in \cite{LIGOScientific:2021aug} ($H_0 = 50^{+37}_{-30} \, \mathrm{km/s/Mpc}$), .
Figure \ref{fig:real_result_multi} shows a peak in the marginalized posterior distribution of $H_0$ around $140\, \mathrm{km/s/Mpc}$.
This means a significant fraction of the reconstructed observed population distribution has an optimized $H_0$ around $140\, \mathrm{km/s/Mpc}$, which does not include the uncertainty from the mismatch between the population model and the observed population distribution.
The reason for the peak differing from the one reported in \cite{LIGOScientific:2021aug}, which is around $50\, \mathrm{km/s/Mpc}$, .
